---
title: "Predicting Mayoral election results in London"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

### Predicting the Mayor of London 2016 results using ward level demographics

In this notebook Multiple Logistic regression, will be used to analyse the results of mayoral election in London 2016 at ward level. The dataset included as **outcome** the results of **Mayor of London election in 2016** and as features (predictors/independent variables) some selected socio-demographic predictors aggregated at ward level.  
I previously used the demographic dataset in another Python projects using Jupyter notebooks: 

**__[Predicting the median house price in London Wards](https://github.com/sebastianBIanalytics/Data_Science_Machine_Learning_Python/blob/master/Predicting%20median%20House%20Price%20London%20-%20Multiple%20Regression.ipynb)__**

**__[Where in London to open a new Luxury Wine Bar](https://github.com/sebastianBIanalytics/Data_Science_Machine_Learning_Python/blob/master/WINE%20BAR%20in%20London%20Final.ipynb)__**



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library (tidyverse)
library (ggplot2)
library(dplyr)
library(plyr)
library(Hmisc)
library(effects)
library(MASS)
library(kableExtra)
library(pROC)
library(fmsb)
library(nlme)
library(rms)

```

The original source of the predictors can be accessed at the link below:  
** Ward Profiles and Atlas of Greater London Authority (GLA)** provided by **__[DataStore London](https://data.london.gov.uk/dataset/ward-profiles-and-atlas)__**. Although the provided details reflect London profile until 2015, this is the most comprehensive dataset publicly available that cover data from the 2011 Census, from ONS and governmental data. 

The election results used the **__[London Elections Results 2016, Wards, Boroughs, Constituency]( https://data.london.gov.uk/download/london-elections-results-2016-wards-boroughs-constituency/01f4ff3a-c562-4d61-977f-c2dfb36694ce/gla-elections-votes-all-2016.xlsx)__**. However, only the voted per ward level were included while the postal votes (given at Borough level) were excluded. 



```{r dataset, include=FALSE}

Election <- read.csv("C:/@@ Default Folder Python Notebooks/Data/Election_London 2016 Features.csv")
Election$Winner <- as.factor(Election$Winner)

```

```{r}
Election$Winner <- revalue(Election$Winner, c("Sadiq Aman Khan"="1","Zac Goldsmith"="0"))
Election$Winner <- factor(Election$Winner, levels = c(0, 1), 
                          labels = c("Zac Goldsmith", "Sadiq Aman Khan"))

```



xxx

```{r}
# Explore the dependent variable

table(Election$Winner)
round(prop.table(table(Election$Winner)),2)

```


```{r}

```


$logit(\text{Vote for S Khan}) = b_0 + b_1\text{Votes for Khan}_i + b_2\text{Median Age}_i +  b_3\text{% Employment 16-64}_i +  b_4\text{Council Tax Bands F/G/H}_i +  b_5\text{Vote Turnout 2012}_i + \varepsilon_i$

Equation (2): Binary logistic regression model equation. 


$$\\[0.5cm]$$

```{r, echo=FALSE}
model1 <- glm(Winner ~ Median.Age.2013 
                      + In.employment..16.64..2011+ X..Households.Private.Rented.2011 , 
                      data = Election, 
                      family = binomial(), na.action = na.fail)

#summary(model1)
```


```{r, echo=FALSE}
election_model <- glm(Winner ~ Median.Age.2013 
                      + In.employment..16.64..2011 
                      + X..Households.Private.Rented.2011 
                      + X..dwellings.CT.bands.F..G.or.H...2015 
                      + Turnout.at.Mayoral.election.2012 + Sadiq.Aman.Khan, 
                      data = Election, 
                      family = binomial(), na.action = na.fail)

#summary(election_model)
```

#### Regression model. 
$$\\[0.1cm]$$
```{r results = "asis"}
stargazer::stargazer(model1, election_model, type = "html", 
title = "Table 1. Multiple Logistic Regression Model", dep.var.labels = "Elected Mayor of London", 
t.auto = FALSE, covariate.labels = c("Median Age 2013", "% In employment 16-64", "Households Private Rented", "Dwellings in CT.bands F/G/H", "Turnout at Mayoral Election 2012", "Sadiq Aman Khan", "Constant"))
```

```{r}

# 1. run a null model 
null_model <- glm(Winner ~ 1, data = Election, family = binomial(), na.action = na.fail) 

# 2. check 
#summary(null_model) 
```


```{r}

# 3. calculate McFadden's R-square 
R2_m1 <- 1-logLik(model1)/logLik(null_model) 

# 4. print it 
R2_m1 
```



```{r}

# 3. calculate McFadden's R-square 
R2 <- 1-logLik(election_model)/logLik(null_model) 

# 4. print it 
R2 
```


```{r}
jtools::effect_plot(election_model, pred = Median.Age.2013, interval = TRUE, 
y.label = "Elected Mayor", x.label = "Median Age in 2013")
```


```{r}
jtools::effect_plot(election_model, pred = In.employment..16.64..2011, interval = TRUE, 
y.label = "Elected Mayor", x.label = "% in Employment 16 - 64")
```


```{r}
jtools::effect_plot(election_model, pred = Sadiq.Aman.Khan, interval = TRUE, 
y.label = "Elected Mayor", x.label = "Votes for Sadiq Khan at ward level")
```


```{r}
jtools::effect_plot(election_model, pred = Turnout.at.Mayoral.election.2012, interval = TRUE, y.label = "Probability election of Sadiq Khan in 2016", x.label = "Turnout Mayoral election 2012")
```

```{r}
jtools::effect_plot(election_model, pred = X..Households.Private.Rented.2011, interval = TRUE, y.label = "Probability election of Sadiq Khan in 2016", x.label = "Households Private Rented 2011")
```


```{r}
jtools::effect_plot(election_model, pred = X..dwellings.CT.bands.F..G.or.H...2015, interval = TRUE, y.label = "Probability election of Sadiq Khan in 2016", x.label = "Dwellings with Council tax in bands F, G or H 2015")
```


```{r}
b <- coef(election_model) %>% exp()  %>%  as_tibble()
b_ci <- suppressMessages(confint(election_model)) %>%  exp() %>%   as_tibble()
OR <- b %>%  bind_cols(b_ci) 

```




```{r}
OR1 <- OR %>% add_column("Labels" = c("Constant", "Median Age 2013", "% In employment 16-64", 
                                      "Households Private Rented", "Dwellings in CT.bands F/G/H", 
                                      "Turnout at Mayoral Election 2012", "Sadiq Aman Khan"), .before = "value")

```


```{r}
kable(OR1, booktabs = TRUE, digits = 3) %>% kable_styling(font_size = 14)

```

```{r}
OR <- data.frame(OR)
```


```{r}

# Estimate the donation probability

Election$Winner_prob1 <- predict(model1, type = "response")
# Find the donation probability of the average prospect
Election$Winner_pred1 <- ifelse(Election$Winner_prob1 > 0.50, 1, 0)

```


```{r}

# Estimate the donation probability

Election$Winner_prob <- predict(election_model, type = "response")
# Find the donation probability of the average prospect
Election$Winner_pred <- ifelse(Election$Winner_prob > 0.50, 1, 0)

```

  

```{r , include=FALSE}
# Create a ROC curve
#ROC1 <- roc(Election$Winner, Election$Winner_prob1)
#ROC <- roc(Election$Winner, Election$Winner_prob)
# Plot the ROC curve
#plot(ROC1, col = "blue")
#plot(ROC, col = "red")
# Calculate the area under the curve (AUC)
#auc(ROC1)
#auc(ROC)
```


```{r, include=FALSE}
summary(model1$fitted.values)
summary(election_model$fitted.values)
```


```{r}

```


```{r, echo=FALSE}

CorelMatrixDataset <- subset(Election, select = -c(ID, Borough, Ward, Constituency, Winner_prob1, Winner_pred1, Winner_prob, Winner_pred))

library(corrgram)
corrgram(CorelMatrixDataset, order=NULL, lower.panel=panel.shade,
  upper.panel=NULL, text.panel=panel.txt,
  main="Correlation matrix")



```


```{r}
# Execute the following command
p1 <- ggplot(Election, aes(Turnout, Sadiq.Aman.Khan)) +
  geom_point()
p1

p2 <- ggplot(Election, aes(Turnout, Sadiq.Aman.Khan, color = Borough)) +
  geom_point()
p2
```
